Rigidity of Expectations: Additional Evidence from Density Forecasts of Professionals and Households

### 2019-07-08 paper/Draft.tex
replace all charts with Core PCE and Core CPI for SPF

	It is widely documented in literature \footnote{The list literature showing positive inflation and uncertainty is long. For example \citet{ball1990inflation}. } and resonated by anecdotal narratives that high inflation is typically associated with high inflation uncertainty. However, the observable correlation of realized inflation and the directly estimated average uncertainty from both professionals and households are at most weakly positive during the period between 2007-2019. In particular, the correlation coefficients  0.14, -0.05 and -0.24 for SPF's CPI forecast, SPF's PCE forecast and SCE's CPI expectation. This may suggest that during a period of persistently low and stable inflation, the conventional positive redux of inflation and uncertainty is not a good description of the relation of the two.

	The middle row of Figure \ref{UnceratitnyOtherMoments} looks into the relationship between size of the forecast error and uncertainty. Although according to our benchmark framework in \ref{theory}, there is no mathematical correlation between the size of the forecast errors and uncertainty as the former depends on the realized shocks and the later depends on the volatility of the shock, it is worth checking if in the data a greater ex ante uncertainty implies bigger ex post forecast errors. The correlation coefficients of the two are -0.19, -0.35 and 0.27 for SPF CPI forecasts, SPF PCE forecasts and SCE's forecasts, respectively. The two indicators turn out to be negatively correlated for professional forecasts. Only households forecasts exhibit such a positive correlation.

	The last row of Figure \ref{UnceratitnyOtherMoments} examines the relationship between disagreement and uncertainty. Many empirical literatures in macroeconomics use disagreements and uncertainty as if they are similar concepts. Such a confusion in practice is partly due to the difficulty of finding appropriate measures of uncertainty in the first place. But as we have presented in \ref{theory}, the two are concepts with distinct statistical definitions. The empirical pattern of the two,  as shown in the plots, also confirms that the two are different objects. The correlation coefficients of the two turn out to be negative, -0.29 and -0.46 for SPF forecasts. In stark contrast, households seem to bear a strong correlation of 0.43 between the two moments.

@@

	\begin{figure}[ht]
		\centering
		\begin{subfigure}[b]{\textwidth}
			\centering
			\caption{Realized Inflation and Uncertainty}
			\label{InfVar}
			\includegraphics[width=0.3\textwidth]{figures/Inf1yf_CPICore_varSPFCPIQ.png}
			\includegraphics[width=0.3\textwidth]{figures/Inf1yf_PCE_varSPFPCEQ.png}
			\includegraphics[width=0.3\textwidth]{figures/Inf1yf_CPIAU_varSCEM.png}
		\end{subfigure}
	    \vspace{1em}
		\vfill
		\begin{subfigure}[b]{\textwidth}
			\centering
			\caption{Size of Forecast Errors and Uncertainty}
			\label{FEVar}
			\includegraphics[width=0.3\textwidth]{figures/SPFCPI_abFE_varSPFCPIQ.png}
			\includegraphics[width=0.3\textwidth]{figures/SPFPCE_abFE_varSPFPCEQ.png}
			\includegraphics[width=0.3\textwidth]{figures/SCE_abFE_varSCEM.png}
		\end{subfigure}
	 \vspace{0.5em}
		\vfill
		\begin{subfigure}[b]{\textwidth}
			\centering
			\caption{Disagreement and Uncertainty}
			\label{DisgVar}
			\includegraphics[width=0.3\textwidth]{figures/CPI_disg_varSPFCPIQ.png}
			\includegraphics[width=0.3\textwidth]{figures/PCE_disg_varSPFPCEQ.png}
			\includegraphics[width=0.3\textwidth]{figures/Q9_disg_varSCEM.png}\\
		\end{subfigure}
		\\
		\begin{flushleft}
			{\footnotesize Note:  From left to right: SPF's  forecasts of core CPI, SPF's core PCE, and SCE's forecast of headline CPI. From top to the bottom, uncertainty (red dash) versus realized inflation (blue dot) with correlation coefficient of 0.14, -0.05 and -0.25, respectively; uncertainty (red dash) versus absolute value of forecast errors (blue dot) with correlation coefficient of -0.19, 0.35, 0.27, respectively; uncertainty (red dash) versus  disagreements (blue dot) with correlation coefficient of -0.29, -0.46 and 0.43.}
		\end{flushleft}
		\caption{Uncertainty and Other Moments}
		\label{UnceratitnyOtherMoments}
	\end{figure}

>>>>

	It is widely documented in literature \footnote{The list literature showing positive inflation and uncertainty is long. For example \citet{ball1990inflation}. } and resonated by anecdotal narratives that high inflation is typically associated with high inflation uncertainty. However, the observable correlation of realized inflation and the directly estimated average uncertainty from both professionals and households are at most weakly positive during the period between 2007-2019. In particular, the correlation coefficients  0.14, 0.12 and -0.24 for SPF's CPI forecast, SPF's PCE forecast and SCE's CPI expectation. This may suggest that during a period of persistently low and stable inflation, the conventional positive redux of inflation and uncertainty is not a good description of the relation of the two.

	The middle row of Figure \ref{UnceratitnyOtherMoments} looks into the relationship between size of the forecast error and uncertainty. Although according to our benchmark framework in \ref{theory}, there is no mathematical correlation between the size of the forecast errors and uncertainty as the former depends on the realized shocks and the later depends on the volatility of the shock, it is worth checking if in the data a greater ex ante uncertainty implies bigger ex post forecast errors. The correlation coefficients of the two are -0.19, -0.18 and 0.27 for SPF CPI forecasts, SPF PCE forecasts and SCE's forecasts, respectively. The two indicators turn out to be negatively correlated for professional forecasts. Only households forecasts exhibit such a positive correlation.

	The last row of Figure \ref{UnceratitnyOtherMoments} examines the relationship between disagreement and uncertainty. Many empirical literatures in macroeconomics use disagreements and uncertainty as if they are similar concepts. Such a confusion in practice is partly due to the difficulty of finding appropriate measures of uncertainty in the first place. But as we have presented in \ref{theory}, the two are concepts with distinct statistical definitions. The empirical pattern of the two,  as shown in the plots, also confirms that the two are different objects. The correlation coefficients of the two turn out to be negative, -0.31 and -0.37 for SPF forecasts. In stark contrast, households seem to bear a strong correlation of 0.43 between the two moments.

@@

	\begin{figure}[ht]
		\centering
		\begin{subfigure}[b]{\textwidth}
			\centering
			\caption{Realized Inflation and Uncertainty}
			\label{InfVar}
			\includegraphics[width=0.3\textwidth]{figures/Inf1yf_CPICore_varSPFCPIQ.png}
			\includegraphics[width=0.3\textwidth]{figures/Inf1yf_PCECore_varSPFPCEQ.png}
			\includegraphics[width=0.3\textwidth]{figures/Inf1yf_CPIAU_varSCEM.png}
		\end{subfigure}
	    \vspace{1em}
		\vfill
		\begin{subfigure}[b]{\textwidth}
			\centering
			\caption{Size of Forecast Errors and Uncertainty}
			\label{FEVar}
			\includegraphics[width=0.3\textwidth]{figures/SPFCPI_abFE_varSPFCPIQ.png}
			\includegraphics[width=0.3\textwidth]{figures/SPFPCE_abFE_varSPFPCEQ.png}
			\includegraphics[width=0.3\textwidth]{figures/SCE_abFE_varSCEM.png}
		\end{subfigure}
	 \vspace{0.5em}
		\vfill
		\begin{subfigure}[b]{\textwidth}
			\centering
			\caption{Disagreement and Uncertainty}
			\label{DisgVar}
			\includegraphics[width=0.3\textwidth]{figures/CORECPI_disg_varSPFCPIQ.png}
			\includegraphics[width=0.3\textwidth]{figures/COREPCE_disg_varSPFPCEQ.png}
			\includegraphics[width=0.3\textwidth]{figures/Q9_disg_varSCEM.png}\\
		\end{subfigure}
		\\
		\begin{flushleft}
			{\footnotesize Note:  From left to right: SPF's  forecasts of core CPI and core PCE, and SCE's forecast of headline CPI. From top to the bottom, uncertainty (red dash) versus realized inflation (blue dot) with correlation coefficient of 0.14, 0.12 and -0.25, respectively; uncertainty (red dash) versus absolute value of forecast errors (blue dot) with correlation coefficient of -0.19, -0.18, 0.27, respectively; uncertainty (red dash) versus  disagreements (blue dot) with correlation coefficient of -0.31, -0.37 and 0.43,repectively. Only the Pearson tests of correlation between disagreement and uncertainty are significant for all.}
		\end{flushleft}
		\caption{Uncertainty and Other Moments}
		\label{UnceratitnyOtherMoments}
	\end{figure}
